---
title: "About Class"
output: 
  html_document:
    toc: true
    toc_depth: 2
    toc_float: true
    theme: cosmo
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE)
library(kableExtra)
```

**Instructors**:  
Rachel Schwartz  
CBLS 377  
4-5404  
rsschwartz@uri.edu 

Please note that the instructor will do her best to respond promptly between 9-5 on weekdays.

**Office Hours**: TBD   
Please feel free to come by without an appointment during this time. To talk with me during other times please set up a meeting by email.

**Class Days / Time**: T/Th 2-3:15 Coastal Institute Kingston 117A

**Prerequisites**: Junior standing and MTH 131 or instructor permission / Graduate Standing

#### Required texts - available free online

[The Practice of Reproducible Research](https://www.practicereproducibleresearch.org/). Justin Kitzes, Daniel Turek, Fatma Deniz (Eds.) 2017. UC Press.

[Modern Dive](http://moderndive.com/). Chester Ismay and Albert Y. Kim. 2017.

[R for Data Science](http://r4ds.had.co.nz/). Hadley Wickham and Garrett Grolemund. 2016. O’Reilly Media.

Data Science from Scratch. Joel Grus. 2015. O’Reilly Media. Free online via the URI library and Safari Books. Search the URI library for the title and follow the links.

Bioinformatics Data Skills. Vince Buffalo. 2015. O’Reilly Media. Free online via the URI library and Safari Books. Search the URI library for the title and follow the links.

[A Whirlwind Tour of Python](https://github.com/jakevdp/WhirlwindTourOfPython). Jake VanderPlas. 2016. Free online via the URI library and Safari Books. Search the URI library for the title and follow the links.  

[Python Data Science Handbook](https://github.com/jakevdp/PythonDataScienceHandbook). Jake VanderPlas. 2017. O'Reilly Media. Free online via the URI library and Safari Books. Search the URI library for the title and follow the links.  

[Data Science at the Command Line](https://www.datascienceatthecommandline.com). Jeroen Janssens. 2014. O'Reilly Media. Free online via the URI library and Safari Books. Search the URI library for the title and follow the links.  

Effective Computation in Physics. Kathryn Huff. 2015. Free online via the URI library and Safari Books. Search the URI library for the title and follow the links.

#### Useful supplementary reads

Doing Data Science. Rachel Schutt and Cathy O'Neill. 2013. O'Reilly Media.

Big Data and Social Science: A Practical Guide to Methods and Tools. 2016. Ian Foster, Rayid Ghani, Ron S. Jarmin, Frauke Kreuter, and Julia Lane, eds. CRC Press.

On Being a Data Skeptic. Cathy O'Neil. 2013. O'Reilly Media.

#### Other equipment / material requirements

Mac, Linux, or Windows laptop (not a tablet, Chromebook, etc.) with administrative privileges. Please contact the instructor if you do not have a laptop and purchasing one would be a financial difficulty.

**Other equipment / material requirements**:

### Assignments and Grading Policy

There will be four assignments during the course of the semester.  Each assignment will focus on the skill learned in class during the week and build on skills and concepts learned in previous weeks. These assignments will be initiated in class and are due at the end of the following week. 

For each assignment you will earn points. The more points you have earned, the more you have learned. Your grade will be based on the total number of points you accumulate during the semester. Assignments will be worth a total of 500 points. 

Students in BIO/DSP539 will also do a project (worth 500 points) involving:  
1. Exploratory analysis of research data (submitted as an Rmarkdown or Jupyter notebook and including explanations)  
2. A ~1000 word paper on your research including relevant and significant results from the analysis.  
3. A Github repository with a Readme so that the analysis can be replicated in full.


```{r, warning=FALSE}
Grade <- c("A","A-","B+","B","B-","C+","C","C-","D+","D","F")
Points_539 <- c("900-1000","880-899","860-879","800-859","780-799","760-779",
             "700-759","680-699","660-679","600-659","<600")
Points_439 <- c("450-500","440-449","430-439","400-429","390-399","380-389",
             "350-379","340-349","330-339","300-329","<300")

grade_539 <- cbind(Grade, Points_539,Points_439)

kable(grade_539, "html") %>%  
  kable_styling(bootstrap_options = c("striped", "hover", "responsive", 
                                      full_width = F))
```

**Special notes**:
If you, or anyone you know, is at risk of going hungry please visit the Rhody Outpost, a food pantry operated by URI. Rhody Outpost is open twice per month at St. Augustine Church at 15 Lower College Rd. For open hours see https://web.uri.edu/deanofstudents/rhody-outpost/ . Rhody Outpost protects the privacy of all visitors. Students may also visit the Jonnycake Center in Peace Dale.

Counseling and Psychological Services: 
Graduate school can be complicated and challenging. You might feel overwhelmed, experience anxiety or depression, or struggle with relationships or family responsibilities. The [URI Counseling Center](https://web.uri.edu/counseling/) provides confidential support for the personal, emotional, and academic well-being of our students. These services include individual counseling, group counseling, and referrals.